The Transition from Massive Metal Free Stars to Dwarf Galaxies

### 2011-01-23 ms.tex
Split M/L and mass fn. figure.  Added Pop III SF section as the first results section.

We initialize the simulation with \textsc{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP $\Lambda$CDM+SZ+LENS best fit
\citep{WMAP7}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b
= 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$ with the
variables having their usual definitions.  We use a maximum refinement
level of $l = 12$, resulting in a maximal comoving resolution of 1 pc.
We stop the simulation at $z=7$ to avoid any large-scale modes with $r
\sim L_{\rm box}$ entering the non-linear growth regime.  At the final
redshift, this simulation has $1.4 \times 10^8$ computational cells.

@@

\begin{enumerate}
\item Radiative cooling from metal species.  We calculate a cooling
  curve with CLOUDY \citep{CLOUDY} and use the method of
  \citep{2008MNRAS.385.1443S}.  Solar abundances were used in the
  cooling curve.  It is tabulated based on density, temperature,
  electron fraction, and total metallicity.
\item A critical metallicity of [Z/H] = --5 and --6, accounting for
  the case where dust cooling can induce low-mass fragmentation in
  these lower metallicities \citep{Omukai05, Schneider06_Frag}.
\item Pop III IMF characteristic mass $M_{\rm char}$ = 40 \Ms.  Using
  the abundance patterns of EMP stars and reionization as constaints,
  the data are best fit with IMFs favoring stars with masses between
  20--130 \Ms \citep{Umeda03, Tumlinson06}.
\item Lyman-Werner radiation background.  We use the semi-analytical
  model of \citet{Wise05}, updated with the 7-year WMAP parameters and
  optical depth to Thompson scattering, to calculate the LW background
  intensity, plotted in Fig. \ref{fig:uvb}.  In this semi-analytical
  model, we use a Pop III stellar mass of 100 \Ms, star formation
  efficiency of 0.005, and escape fraction of 0.2 \citep{Wise09}.  The
  LW radiation from point sources are added to the background.  The
  intensity decreases after $z \sim 14$ because Pop II star formation
  becomes dominant, which produce less LW specific luminosity than Pop
  III stars that have surface temperatures of $T = 10^5$ K.  For
  computational convenience, we fit the background evolution with the
  function
%
  \begin{equation}
    \label{eqn:uvb}
    \log J_{21}(z) = A + Bz - Cz^2 + Dz^3 - Ez^4,
  \end{equation}
%
  where $(A,B,C,D,E) = (-2.356, 0.4562, 0.02680, 5.882 \times 10^{-4},
  -5.056 \times 10^{-6})$ and $J_{21}$ is the specific intensity in
  units of $10^{-21}$ \emis.  This fits the
  model data within 1\% in $6 \le z \le 30$ and is consistent with
  $J_{21}$ values in \citet{Trenti09_SFR}.
\item Pop II star formation only.  This model is useful to study the
  importance of treating Pop III star formation as single stars and
  their pair-instability and Type II SNe.
\item Pop II star formation only and no molecular hydrogen cooling.
  This model provides a good comparison to larger scale simulations
  where it is not computationally feasible to consider \hh~cooling or
  resolve the necessary scales for Pop III star formation.
\end{enumerate}

@@

\begin{figure}
  \plotone{Lstar.eps}
  \caption{\label{fig:mass_fn} The top panel shows the halo mass
    function of the simulated haloes (points) where the error bars are
    Poisson noise and the solid line is the analytic halo mass
    function of \citet{Warren06}.  The bottom panel depicts the galaxy
    luminosities versus halo DM mass.  Overplotted are constant
    mass-to-light ratios.  The large scatter at low halo masses are
    similar to ones observed in local dSph galaxies.}
\end{figure}

@@

Here we present the gaseous and stellar evolution of the high-redshift
dwarf galaxies that form in our simulations.  To illustrate its global
properties, we show the density-weighted projections of gas density,
temperature, and metallicity at redshifts 15, 12, 10, 8, and 7 in
Fig. \ref{fig:evo-mosaic}.  Most apparent is the photo-heating of
IGM over this redshift range, where the ionized fraction is 0.XX at
$z=7$.  The metallicity projections show the growth of metal-enriched
bubbles, starting first with enrichment from Pop III stars.  Then as
time progresses, the haloes enriched within these bubbles above $Z_{\rm
  cr}$ transition to Pop II star formation, enriching their host halo
and surrounding IGM.  In Fig. \ref{fig:mass_fn}, we show the halo mass
function, agreeing well with the analytic halo mass function of
\citet{Warren06}.  We identify the haloes with a friends-of-friends
(FOF) algorithm \citep{Davis85}.  The excess of low-mass haloes below
$M_{\rm DM} = 2 \times 10^6 \Ms$ is most likely associated with the
biased region in the lower-right region in the projections.
Furthermore, the uncertainities in different analytic halo mass
functions \citep[e.g., see][for a comparison]{Reed07} and halo
identification are comparable than these differences.  Also in this
Figure, we display the galaxy luminosities with respect to halo DM
mass found with the FOF halo finder.  Our fidicual model contains 38
galaxies with a total of 3640 Pop II star clusters and captures the
formation of 333 Pop III stars.  Note that the most massive halo
($M_{\rm DM} = 5.2 \times 10^8 \Ms$) is undergoing a starburst at
$z=7$ and has a high $M/L = 3$.  The lower mass galaxies show a spread
in M/L over two orders of magnitude, similar to the observations of
local dSph galaxies \citep[e.g.][]{Strigari08}.

We split our presentation into two sets of analyses.  In one, we focus
on the evolution of the individual halo and galaxy properties of the
10 most massive haloes at $z=7$ to better understand their formation
process and progenitors.  We list the global properties of these
haloes at $z = 7$ in Table \ref{tab:halos}.  Second, we analyze the
statistics of the star formation and metal enrichment in all 38
galaxies.  We start by presenting the change in the global halo
properties, followed by a description of their star formation
histories, metallicity distribution functions, and star formation
rates.  We then present their metal enrichment and lastly comment on
unique details of individual haloes.

@@

@@

@@

\subsubsection{Metallicity distribution function}

@@

\begin{figure*}
  \epsscale{2.0}
  \plotone{sfh1.eps}
  \epsscale{1.0}
  \caption{\label{fig:SFR} Star formation history.}
\end{figure*}

\subsection{Metal enrichment}

\begin{figure}
  \plotone{Zhistory.eps}
  \caption{\label{fig:Zevo} Metal evolution.}
\end{figure}

\begin{figure}
  \plottwo{z2.eps}{z3.eps}
  \caption{\label{fig:zhalo} Halo metallicities.}
\end{figure}

>>>>

We initialize the simulation with \textsc{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP $\Lambda$CDM+SZ+LENS best fit
\citep{WMAP7}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b
= 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$ with the
variables having their usual definitions.  We use a maximum refinement
level of $l = 12$, resulting in a maximal comoving resolution of 1 pc.
We stop the simulation at $z=7$ to avoid any large-scale modes with $r
\sim L_{\rm box}/2$ entering the non-linear growth regime.  At the
final redshift, this simulation has $1.4 \times 10^8$ computational
cells.

@@

\begin{enumerate}
\item Radiative cooling from metal species.  We calculate a cooling
  curve with CLOUDY \citep{CLOUDY} and use the method of
  \citep{2008MNRAS.385.1443S}.  Solar abundances were used in the
  cooling curve.  It is tabulated based on density, temperature,
  electron fraction, and total metallicity.
\item A critical metallicity of [Z/H] = --5 and --6, accounting for
  the case where dust cooling can induce low-mass fragmentation in
  these lower metallicities \citep{Omukai05, Schneider06_Frag}.
\item Pop III IMF characteristic mass $M_{\rm char}$ = 40 \Ms.  Using
  the abundance patterns of EMP stars and reionization as constaints,
  the data are best fit with IMFs favoring stars with masses between
  20--130 \Ms~\citep{Umeda03, Tumlinson06}.
\item Lyman-Werner radiation background.  We use the semi-analytical
  model of \citet{Wise05}, updated with the 7-year WMAP parameters and
  optical depth to Thompson scattering, to calculate the LW background
  intensity, plotted in Fig. \ref{fig:uvb}.  In this semi-analytical
  model, we use a Pop III stellar mass of 100 \Ms, star formation
  efficiency of 0.005, and escape fraction of 0.2 \citep{Wise09}.  The
  LW radiation from point sources are added to the background.  The
  intensity decreases after $z \sim 14$ because Pop II star formation
  becomes dominant, which produce less LW specific luminosity than Pop
  III stars that have surface temperatures of $T = 10^5$ K.  For
  computational convenience, we fit the background evolution with the
  function
%
  \begin{equation}
    \label{eqn:uvb}
    \log J_{21}(z) = A + Bz - Cz^2 + Dz^3 - Ez^4,
  \end{equation}
%
  where $(A,B,C,D,E) = (-2.356, 0.4562, 0.02680, 5.882 \times 10^{-4},
  -5.056 \times 10^{-6})$ and $J_{21}$ is the specific intensity in
  units of $10^{-21}$ \emis.  This fits the
  model data within 1\% in $6 \le z \le 30$ and is consistent with
  $J_{21}$ values in \citet{Trenti09_SFR}.
\item Pop II star formation only.  This model is useful to study the
  importance of treating Pop III star formation as single stars and
  their pair-instability and Type II SNe.
\item Pop II star formation only and no molecular hydrogen cooling.
  This model provides a good comparison to larger scale simulations
  where it is not computationally feasible to consider \hh~cooling or
  resolve the necessary scales for Pop III star formation.
\end{enumerate}

@@

\begin{figure}
  \plotone{mass_fn.eps}
  \caption{\label{fig:mass_fn} Halo DM mass function of the simulated
    haloes (points) where the error bars are Poisson noise and the
    solid line is the analytic halo mass function of
    \citet{Warren06}.}
\end{figure}

@@

Here we present the gaseous and stellar evolution of the high-redshift
dwarf galaxies that form in our simulations.  To illustrate its global
properties, we show the density-weighted projections of gas density,
temperature, and metallicity at redshifts 15, 12, 10, 8, and 7 in
Fig. \ref{fig:evo-mosaic}.  Most apparent is the photo-heating of IGM
over this redshift range, where the ionized fraction is 0.XX at $z=7$.
The metallicity projections show the growth of metal-enriched bubbles,
starting first with enrichment from Pop III stars.  Then as time
progresses, the haloes enriched within these bubbles above $Z_{\rm
  cr}$ transition to Pop II star formation, enriching their host halo
and surrounding IGM.  In Fig. \ref{fig:mass_fn}, we show the halo mass
function, agreeing well with the analytic halo mass function of
\citet{Warren06}.  We identify the haloes with a friends-of-friends
(FOF) algorithm \citep{Davis85}.  The excess of low-mass haloes below
$M_{\rm DM} = 2 \times 10^6 \Ms$ is most likely associated with the
biased region in the lower-right region in the projections.
Furthermore, the uncertainities in different analytic halo mass
functions \citep[see][for a comparison]{Reed07} and halo
identification are comparable than these differences.  Our fidicual
model contains 38 galaxies with a total of 3640 Pop II star clusters
and captures the formation of 333 Pop III stars.

We split our presentation into two sets of analyses.  In one, we focus
on the evolution of the individual halo and galaxy properties of the
10 most massive haloes at $z=7$ to better understand their formation
process and progenitors.  We list the global properties of these
haloes at $z = 7$ in Table \ref{tab:halos}.  Second, we analyze the
statistics of the star formation and metal enrichment in all 38
galaxies.  We first present results on the frequency of Pop III star
formation and the chemical enrichment from their PISNe.  We then focus
on the evolution in the global halo properties, followed by a
description of their Pop II star formation histories, metallicity
distribution functions, and star formation rates.  We then present
their metal enrichment and lastly comment on unique details of
individual haloes.

\subsection{Pop III star formation}

\begin{figure}
  \plotone{z3.eps}
  \caption{\label{fig:zhalo3} Average halo metallicity from Pop III
    SNe as a function of total halo mass.  Inhomogenous metal
    enrichment causes the large spread in metallicity at low masses.
    At $M_{\rm halo} > 3 \times 10^7 \Ms$, all halos are enriched by
    Pop III SNe above [Z/H] $>$ --4.  The y-axis error bars show the
    25\% and 75\% percentile values because the underlying
    distribution is not Gaussian.}
\end{figure}

\begin{figure}
  \plotone{Zhistory.eps}
  \caption{\label{fig:Zevo} The time evolution of the average gas
    metallicity from Pop II SNe (dashed) and Pop III SNe (dotted) and
    average stellar metallicity (solid) of the 10 most massive haloes.
    Note that a metallicity floor of [Z/H] $\sim -3$ exists in every
    halo.}
\end{figure}

\subsubsection{Number of Pop III stars}

In Table \ref{tab:halos} we list the number of the Pop III stars,
$N_{\rm III}$, that have formed in the progenitors of the 10 most
massive haloes.  For Haloes 1--9, between one and four Pop III stars
formed in the progenitors.  The progenitors of Halo 0 hosted 21 Pop
III stars, which is consistent with the number of Pop III stars per
$10^8 \Ms$ halo seen in the lower mass haloes, considering that
several such haloes merged to form the $10^9 \Ms$ halo.

Given the idea of most minihaloes host a Pop III, these numbers appear
low.  However there are a few processes that can suppress Pop III star
formation:
%
\begin{enumerate}
\item photo-dissociation of \hh~\citep{Dekel87, Haiman00, Machacek01},
\item photo-evaporation of minihaloes \citep{Shapiro04, whalen08}, and
\item metal enrichment from nearby (PI)SNe \citep{Wise08_Gal, Greif10,
  Trenti09}.
\end{enumerate}
We observe all of these affecting Pop III star formation in the
simulation, confirming that negative feedback plays an important role
in metal-free star formation.

\subsubsection{Metal enrichment from pair-instability SNe}

@@

\begin{figure}
  \vspace{200pt}
  \caption{\label{fig:filtering} Halo filtering mass.}
\end{figure}

@@

\subsection{Pop II star formation}

\begin{figure*}
  \epsscale{2.0}
  \plotone{sfh1.eps}
  \epsscale{1.0}
  \caption{\label{fig:SFH} The scatter plots show the Pop II star
    formation history as a function of metallicity for the 10 most
    massive halos.  Halo 8 is not shown because it contains no stars.
    Each circle represents a star particle, whose area is proportional
    to its mass.  The open circles in the upper right represent $M =
    10^3$, $10^4$, and $10^5$ \Ms.  The dashed lines in the Halo 0
    panel guide the eye to two stellar populations that were formed in
    two satellite halos, merging at $z=7.5$.  The upper histogram
    shows the star formation rate.  The right histogram depicts the
    Pop II metallicity distribution function.}
\end{figure*}

\begin{figure}
  \plotone{Lstar.eps}
  \caption{\label{fig:Lstar} Galaxy luminosities versus total halo
    mass.  Overplotted are constant mass-to-light ratios.  The large
    scatter at low halo masses are similar to ones observed in local
    dSph galaxies.}
\end{figure}

@@

The scatter plots of Fig. \ref{fig:SFH} show the Pop II star formation
history as a function of metallicity for the 10 most massive halos.

We now focus on the star formation rates (SFR) and metallicity
distribution functions (MDF) of these histories, collapsing the
stellar population in the metallicity and time axes, respectively.

@@

\subsubsection{Metallicity distribution functions}

\subsubsection{Mass-to-light ratios}
\label{sec:mlratios}

Note that the most massive halo ($M_{\rm tot} = 10^9 \Ms$) is
undergoing a starburst at $z=7$ and has a relatively high $M/L = 6$.
The lower mass galaxies show a spread in M/L over two orders of
magnitude, similar to the Sloan Digital Sky Survey (SDSS) observations
of local dSph galaxies \citep[e.g.][]{Strigari08}.

\subsubsection{Metal enrichment}

\begin{figure}
  \plotone{z2.eps}
  \caption{\label{fig:zhalo2} Halo metallicities.}
\end{figure}
